### 2022-02-05 outputs/paper/paper.Rmd
commit for the day

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Arsh Lakhanpal
thanks: "Code and data are available at: 
https://github.com/lakhan99/starter_folder-main"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "The city of Toronto has been able to collect data pertraining to COVID-19 cases demographic and geographic variables since January 2020. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(patchwork)
library(tidyverse)
library(palmerpenguins)
library(opendatatoronto)
library(dplyr)
library(tidygeocoder)
library(lubridate)
library(janitor)
library(ggmap)
covid_cases <- read_csv(here::here("inputs/data/covid_cases.csv"))

```

# Introduction

On December 31, 2019 the world saw its first case of the coronavirus, officially referred to as SARS-CoV-2 or COVID-19, in Wuhan China and by January 25, 2020 Toronto had seen its first case. Ever since, the virus continues to spread in various forms like the Delta variant and the most recent being the Omicron variant. The province has taken many attempts at fighting the virus which include locking down the province and shutting down non-essential activities in hopes of limiting close contact between citizens.

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

```{r, echo=FALSE}
covid_cases <- clean_names(covid_cases)
#covid_cases <- subset(covid_cases, select = -c(1,2,7,8))
```

  The statistical programming language used to work with the data in this paper is R (add citation) along with packages like tidyverse, knitr. The "COVID-19 Cases in Toronto" dataset was obtained from Open Data Toronto which houses a broad number of datasets. This data set was published by Toronto Public Health and for this paper, this dataset was officially imported from the opendatatoronto package in R.

  This data consists of confirmed and probable reported COVID-19 cases among Toronto residents with demographic, geographical and epidemiological variables.(citation) The data itself is claimed to have been extracted from the provincial Case and Contact Management System. It is important to note that as the pandemic is happening, this dataset continues to get updated therefore results of this dataset are based off data up till January 24,2022. The dataset consists of 272017 observations of 18 variables however, since the first two variables were simply just "id" and "Assigned_ID" these were removed as they held no importance the the focus of this paper.

COVID-19, like many other viruses thrives in colder drier climates and (Figure \@ref(fig:month)) the number of cases for every month in 2020 and 2021.  Using the tools from the dplyr package, two new variables named "month_of_ diagnosis" and "year_of_diagnosis" were mutated from the original variable "episode_date" which pertained to the date that the case was confirmed. The following two charts show the number of covid cases for both 2020 and 2021 as those are the only two complete years that we have have to live with the virus.

```{r month, echo=FALSE}

covid_cases <- covid_cases %>% 
  mutate(month_of_diagnosis = month(episode_date,
                  label = TRUE,
                  abbr = TRUE))

covid_cases <- covid_cases %>% mutate(year_of_diagnosis = year(episode_date))
covid_cases2020 <- covid_cases %>% filter(year_of_diagnosis == "2020")
covid_cases2021 <- covid_cases %>% filter(year_of_diagnosis == "2021")

fig2020 <- covid_cases2020 %>% ggplot(mapping = aes(x= month_of_diagnosis)) + geom_bar() + labs(x = "Month of 2020", y = "Number of COVID cases")

fig2021 <- covid_cases2021 %>% ggplot(mapping = aes(x= month_of_diagnosis)) + geom_bar() + labs(x = "Months of 2021", y = "Number of COVID cases")

monthly_figs <- (fig2020 / fig2021)
monthly_figs
monthly_figs + plot_annotation(caption = "Figure 1: COVID-19 Cases for every month in 2020 and 2021", theme = theme(plot.caption.position = "plot", plot.caption = element_text(hjust = 0.1)))

```
From Figure \@ref(fig:month), the colder months in the year tend to have the most number of cases where as the summer months have a significantly lower number of cases. In 2020, we can see how the number of covid cases were highest during the fall and winter months and how that trend carries onto early 2021 which should be what is expected. The number of cases increased near the winter 2020 and that trend continued in until April 2021 where the cases began to continuously decrease. Unlike 2021, where we saw an increasing trend in the colder months, for 2021 it seems that the number of cases had stabilized, possibly due to the increased amount of the population that the vaccine had been administered to by then. As part of Ontario's "Phase 2," vaccines had begun to be distributed based off "hot spots" by April of 2021 and eligible to everyone as a part of their "Phase 3" which began in July of 2021.

One of the possible reasons explaining this rise in cases during this time is the fact that this is around the holidays when families tend to get together.

```{r tableone, echo=FALSE}

S <- count(covid_cases2020 %>% 
  filter(month_of_diagnosis == "Sep", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

O <- count(covid_cases2020 %>% 
  filter(month_of_diagnosis == "Oct", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

N <- count(covid_cases2020 %>% 
  filter(month_of_diagnosis == "Nov", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

D <- count(covid_cases2020 %>% 
  filter(month_of_diagnosis == "Dec", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

S1 <- count(covid_cases2021 %>% 
  filter(month_of_diagnosis == "Sep", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

O1 <- count(covid_cases2021 %>% 
  filter(month_of_diagnosis == "Oct", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

N1 <- count(covid_cases2021 %>% 
  filter(month_of_diagnosis == "Nov", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

D1 <- count(covid_cases2021 %>% 
  filter(month_of_diagnosis == "Dec", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

September <- c(3012,2806)
October <- c(2413,1429)
November <- c(5029, 1932)
December <- c(9047, 3975)

holiday_case <- data.frame(September, October, November, December)
rownames(holiday_case) <- c("2020","2021")

holiday_case<- knitr::kable(holiday_case, caption = "Number of covid cases from contact during holiday months")
holiday_case
```

Table \@ref(tab:tableone) shows the count for COVID-19 cases that were transmitted through contact which includes close contact, household contact and community contact. For both years, we notice that the month of December had the highest number of COVID-19 cases, likely because of Christmas. Another thing to consider is that in December of 2020, there was a new and more contagious variant of the COVID-19 virus called the delta variant. Similarly, at the end of November 2021, we got the omicron variant, one that was even more contagions than the delta variant which also explains the high case counts in December of 2020 and 2021. The vaccines that people had taken until December of 2021 did not necessarily protect them from the omicron variant which is likely to be another reason that December of 2021 had the largest amount of cases for any month.

Another variable this data set gave us was the age group that every case fell under. These age groups ranged from "19 and younger" to "90 and older." The following figure shows the number of cases for every age group.

```{r, echo=FALSE}
age_cases <- covid_cases %>% ggplot(mapping = aes(x= age_group)) + geom_bar() + labs(x = "Age Group", y = "Number of COVID-19 cases") + coord_flip()

age_cases + plot_annotation(caption = "Cases of COVID-19 by age group", theme = theme(plot.caption.position = "plot", plot.caption = element_text(hjust = 0.1)))

```
Figure 3 (change later) shows us that the younger population catches the virus much more compared to the older population, likely because the younger population has more of a reason to be outside the house, whether that be for work or for school. This could also explain the reasoning behind the decisions of closing schools, restaurants, bars and other recreational facilities since these are all places where one is more likely to find a younger audience. Figure 4 gives a more holistic view of the number of covid cases as it incorporates the age group and number of covid cases per month into one graph. For both years, it is evident that those between the ages of 20 and 40 have the highest proportions of cases regardless of which month is looked at.

shows us  
(talk about possible lockdown/vaccine trends)

We can further focus on the covid cases for every month and which age group seemed to be most affected by covid throughout this pandemic.

```{r}
ggplot(data = covid_cases2020, mapping = aes_(x = covid_cases2020$month_of_diagnosis, fill = covid_cases2020$age_group)) + 
                         geom_bar(postion = 'stack')

ggplot(data = covid_cases2021, mapping = aes_(x = covid_cases2021$month_of_diagnosis, fill = covid_cases2021$age_group)) + 
                         geom_bar(postion = 'stack')

#covid_cases %>% count(age_group)

"This is used to get a visual representation of the age groups being infected"

#seperate this into 2020,2021 as well?

```
This dataset is made up of data collected from individuals which means that there is likely to be some sort of ethical considerations. One of the largest concerns with this dataset is that it only accounts for the "confirmed" and "probable cases." This leaves out many categories of cases that likely contain a lot of individuals:

*The individual is unknowingly living with COVID
  + Since the virus does affect every individual differently, there are many cases where the individual isn't even aware of the fact that they have the virus and continue to live with it until it naturally goes away.

* The individual knows they have the virus but refuses to report it

* The individual feels sick but mistake it for the common cold/other virus.

#gender reason as well?

```{r}

covid_cases %>% count(client_gender)
covid_cases2021 %>% ggplot(mapping = aes(x= age_group)) + geom_bar()

#fatal_2020 <- covid_cases %>% 
  #filter(year_of_diagnosis == "2020" & outcome == "FATAL") %>% 
  #select(year_of_diagnosis, outcome)

#resolved_2020 <- covid_cases %>% 
  #filter(year_of_diagnosis == "2020" & outcome == "RESOLVED") %>% 
  #select(year_of_diagnosis, outcome)

#fatal_2021 <- covid_cases %>% 
  #filter(year_of_diagnosis == "2021" & outcome == "FATAL") %>% 
  #select(year_of_diagnosis, outcome)

#resolved_2021 <- covid_cases %>% 
  #filter(year_of_diagnosis == "2021" & outcome == "RESOLVED") %>% 
  #select(year_of_diagnosis, outcome)

#covid_cases %>% count(year_of_diagnosis)

#covid_cases %>% count(month_of_diagnosis)
#covid_cases %>% count(outcome)
#propo <- covid_cases %>% count(source_of_infection)

covid_cases %>% count(age_group)
#cases_per_year <- covid_cases %>% select(year_of_diagnosis, outcome) 
#API

#knitr:: kable(table_c, 
              #col.names = "January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
```

```{r}
#locations <- tibble(
  #postal = c(covid_cases$fsa)
  #)

#hopeful <- geocode(locations,
        #postalcode = "postal",
        #method = "osm")

#cities <- revgeocode(c(43.77977

#,-79.36618), output = "address")

#cities$locality

```

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
#ggplot(penguins, aes(x = island, fill = species)) +
 # geom_bar(alpha = 0.8) +
  #scale_fill_manual(values = c("darkorange","purple","cyan4"),
   #                 guide = FALSE) +
  #theme_minimal() +
  #facet_wrap(~species, ncol = 1) +
  #coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?

```

Talk way more about it.

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

\newpage

# References